Modal Verification Patterns for Systems

### 2025-06-25 paper.tex
+
views finalized

%\bibliography{references/paper}
\begin{document}
%
\title{Survey: Modal Verification Patterns for Systems}
%
%\titlerunning{RCU Types}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
%\author{Ismail Kuru \orcidID{0000-0002-5796-2150} \and
%Colin S. Gordon \orcidID{0000-0002-9012-4490}}
%%%Third Author\inst{3}\orcidID{2222--3333-4444-5555}}
%
%\authorrunning{Kuru and Gordon}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
%\institute{Drexel University \\ %\and
%%%Springer Heidelberg, Tiergartenstr. 17, 69121 Heidelberg, Germany
%\email{\{ik335,csgordon\}@drexel.edu}}%\\
%%%\url{http://www.springer.com/gp/computer-science/lncs} \and
%%%ABC Institute, Rupert-Karls-University Heidelberg, Heidelberg, Germany\\
%%%\email{\{abc,lncs\}@uni-heidelberg.de}}
%
\maketitle              % typeset the header of the contribution

@@

\subsubsection{Reference-Counting Memory Reclamation}
In a typical configuration, allocating, deallocating and referencing memory locations for client code 
leaves the memory layout \emph{ideally} in a form of DAG (directed acyclic graph) with a unique root, which resembles the first stack pointer referencing the allocated memory. DAG abstraction instead of recursive decent tree allows non-duplication on the reference counts for the identical memory objects.
% To ease picturing the memory layout after a set of requests from the client code (for allocation, referencing etc.), we can think of the unique roots of DAGs as 
% \emph{super} block addresses -- this time multiple ones, not only two -- in Figure \ref{fig:cowindex}, 
% from which a path is created for a stack pointer reference to the allocated memory. 
\todo[inline,color=cyan]{
Why is this a DAG? Normally programs can easily create cycles, so the DAG bit requires a little more explanation.
Do we mean that it is ``often'' the case that it's a DAG?}

@@

This choice trades off complexity against flexibility and scoping constraints.
Strongly nominalized elements track additional specifier/prover-visible book-keeping data.
But in exchange for carrying those identifiers of their context with them, 
strongly nominalized elements can be used together under any modality. For example, one can
use the StackRegion modality to talk about two different stack frames simultaneously for code
which accesses multiple stack frames: $\fbox{n} \; \ell \mapsto v \ast \fbox{n+1} \; \ell' \mapsto v'$.
Using a given strongly nominalized assertion element under different modalities for different
frames does not change its meaning. Likewise,  weak-memory specific modalities existentially quantify over other views, related to the ``current'' view (the one where
the current thread's assertions are evaluated), and evaluate a resource (an thread-local view of an address-value mapping) with respect to those other views.

>>>>

%\bibliography{references/paper}
\begin{document}
%
\title{Modal Verification Patterns for Systems}
%
%\titlerunning{RCU Types}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
%\author{Ismail Kuru \orcidID{0000-0002-5796-2150} \and
%Colin S. Gordon \orcidID{0000-0002-9012-4490}}
%%%Third Author\inst{3}\orcidID{2222--3333-4444-5555}}
%
%\authorrunning{Kuru and Gordon}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
%\institute{Drexel University \\ %\and
%%%Springer Heidelberg, Tiergartenstr. 17, 69121 Heidelberg, Germany
%\email{\{ik335,csgordon\}@drexel.edu}}%\\
%%%\url{http://www.springer.com/gp/computer-science/lncs} \and
%%%ABC Institute, Rupert-Karls-University Heidelberg, Heidelberg, Germany\\
%%%\email{\{abc,lncs\}@uni-heidelberg.de}}
%
\maketitle              % typeset the header of the contribution

@@

\subsubsection{Reference-Counting Memory Reclamation}
In a typical configuration, allocating, deallocating and referencing memory locations for client code 
leaves the memory layout \emph{ideally} in a form of DAG (directed acyclic graph) with a unique root, which resembles the first stack pointer referencing the allocated memory. DAG abstraction instead of recursive decent tree allows non-duplication on the reference counts for the identical memory objects.
% To ease picturing the memory layout after a set of requests from the client code (for allocation, referencing etc.), we can think of the unique roots of DAGs as 
% \emph{super} block addresses -- this time multiple ones, not only two -- in Figure \ref{fig:cowindex}, 
% from which a path is created for a stack pointer reference to the allocated memory. 
%\todo[inline,color=cyan]{
%Why is this a DAG? Normally programs can easily create cycles, so the DAG bit requires a little more explanation.
%Do we mean that it is ``often'' the case that it's a DAG?}

@@

This choice trades off complexity against flexibility and scoping constraints.
Strongly nominalized elements track additional specifier/prover-visible book-keeping data.
But in exchange for carrying those identifiers of their context with them, 
strongly nominalized elements can be used together under any modality. For example, one can
use the StackRegion modality to talk about two different stack frames simultaneously for code
which accesses multiple stack frames: $\fbox{n} \; \ell \mapsto v \ast \fbox{n+1} \; \ell' \mapsto v'$.
Using a given strongly nominalized assertion element under different modalities for different
frames does not change its meaning. Likewise,  weak-memory specific modalities existentially quantify over other views, related to the ``current'' view (the one where
the current thread's assertions are evaluated), and evaluate a resource (a thread-local view of an address-value mapping) with respect to those other views. Given that views are identified by \emph{nominals}, referring to all the other views makes weak-memory modalities strongly nominalized.
